---
title: "Initial_testing"
author: "Alan Jackson"
format: html
editor: source
---

##    Test reading file and posting on map

```{r}

# setup

library(tidyverse)

path <- "/home/ajackson/Dropbox/Rprojects/TrashTracker/data/"


```

##        Read in all the files

Read files in, and split into path and collection points
Run some simple stats on collection points
Project the data
Create a 25 m corridor around the path
Are there collection points outside that corridor?
If so create 10 meter circles around them, and do a union with corridor
Create a 5 m grid in the area
Using 20 m mean filter, populate grid
Contour and plot


```{r}

#   Read in the files

gpx_parsed <- XML::htmlTreeParse(file = paste0(path, "20240620.gpx"), useInternalNodes = TRUE)

coords <- XML::xpathSApply(doc = gpx_parsed, path = "//trkpt", fun = XML::xmlAttrs)
ways <- XML::xpathSApply(doc = gpx_parsed, path = "//wpt", fun = XML::xmlAttrs)
wpt_type <- XML::xpathSApply(doc = gpx_parsed, path = "//wpt/name", fun = XML::xmlValue)
elevation <- XML::xpathSApply(doc = gpx_parsed, path = "//wpt/ele", fun = XML::xmlValue)
# elevation <- XML::xpathSApply(doc = gpx_parsed, path = "//trkpt/ele", fun = XML::xmlValue)

df <- data.frame(
  lat = as.numeric(ways["lat", ]),
  lon = as.numeric(ways["lon", ]),
  waypt = wpt_type,
  elevation = as.numeric(elevation)
)

```

## plot it

```{r}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addPolylines(data = df, lat = ~lat, lng = ~lon, color = "#000000", opacity = 0.8, weight = 3) %>% 
  addMarkers(data=df, lat = ~lat, lng = ~lon )
```


##        Better plot

```{r}

df2 <- df %>% 
  sf::st_as_sf(., coords=c("lat","lon"),crs=4326)


tmap::tmap_options(basemaps="OpenStreetMap")
tmap::tmap_mode("view") # set mode to interactive plots

  tmap::tm_shape(df2) + 
  tmap::tm_dots(size=0.05) +
  tmap::tm_shape(df2) + 
  tmap::tm_dots(col="red")
```















