---
title: "Read data and clean up"
author: "Alan Jackson"
format: html
editor: source
---

##    Read files, project, and post on map

Read files in, and split into path and collection points
Run some simple stats on collection points
Project the data
Despike
Create a smooth path
Create a 25 m corridor around the path
Are there collection points outside that corridor?
If so create 10 meter circles around them, and do a union with corridor
Create a 5 m grid in the area
Using 20 m mean filter, populate grid
Contour and plot

```{r}

# setup

library(tidyverse)

googlecrs <- 4326
Seattle <- "EPSG:2285"

path <- "/home/ajackson/Dropbox/Rprojects/TrashTracker/dataSeattle/"


```

##    Functions

```{r}


#   distance function

Eucl <- function(x1, x2, y1, y2){sqrt((x1-x2)**2 + (y1-y2)**2)}

#     Delta dist

delta_Dist <- function(df){
  df %>% 
    mutate(X=sf::st_coordinates(.)[,1],
           Y=sf::st_coordinates(.)[,2]) %>% 
    group_by(Date) %>% 
      mutate(Xtest=(lag(X, default=first(X)) + lead(X, default=last(X)))/2) %>% 
      mutate(Ytest=(lag(Y, default=first(Y)) + lead(Y, default=last(Y)))/2) %>% 
      mutate(Dx=Xtest - X,
             Dy=Ytest - Y) %>% 
      mutate(Delta_dist=Eucl(Xtest, X, Ytest, Y)) %>% 
      mutate(New_dist=Eucl(lag(X, default=first(X)), lead(X, default=last(X)), 
                           lag(Y, default=first(Y)), lead(Y, default=last(Y)))) %>% 
    ungroup()
}

###   Kalman filter

# Kalman <- function(day) {
Kalman <- function(X, Y, varM, varP) {
  #initializing variables
  # count <- nrow(day) # number of data points in the day
  count <- length(X) # number of data points in the day
  print(paste("K-1", count))
  # z <- cbind(day$X,day$Y) #measurements
  z <- cbind(X,Y) #measurements
  #Allocate space:
  xhat <- matrix(rep(0,2*count),ncol =2) #a posteri estimate at each step
  P <- array(0,dim=c(2,2,count))  #a posteri error estimate
  xhatminus <- matrix(rep(0,2*count),ncol =2) #a priori estimate
  Pminus <- array(0,dim=c(2,2,count)) #a priori error estimate
  K <- array(0,dim=c(2,2,count)) #gain
  #Initializing matrices
  A <-diag(2)
  H<-diag(2)
  # R<-function(k) diag(2)* day$accuracy[k]^2#estimate of measurement variance
  R<-function(k) diag(2)* varM[k] #estimate of measurement variance
  # Q<-function(k) diag(2)* as.numeric(day$timeNextMeasure[k])^1.5# the process variance
  Q<-function(k) diag(2)* varP[k] # the process variance
  #initialise guesses:
  xhat[1,] <- z[1,]
  P[,,1] <- diag(2)
  for (k in 2:count){
    #time update
    #project state ahead
    xhatminus[k,] <- A %*% xhat[k-1,] #+ B %% u[k-1]
    #project error covariance ahead
    Pminus[,,k] <- A %*% P[,,k-1] %*%  t(A) + (Q(k))
    #measurement update
    # kalman gain
    K[,,k] <- Pminus[,,k] %*% t(H)/ (H %*% Pminus[,,k] %*% t(H) + R(k))
    #what if NaaN?
    K[,,k][which(is.nan(K[,,k]))]<-0
    # update estimate with measurement
    
    xhat[k,] <-  xhatminus[k,] + K[,,k] %*% (z[k,] - H %*% xhatminus[k,])
    #update error covariance
    P[,,k] = (diag(2) - K[,,k]%*% H) %*% Pminus[,,k]
  }
  return(as.data.frame(xhat) %>% rename(X2=V1, Y2=V2))
  # cbind(day, as.data.frame(xhat) %>% rename(X2=V1, Y2=V2))
}

```

##        Read in all the files

Read in files
split
generate stats

```{r}

#   Read in the files

in_files <- list.files( path=paste0(path), pattern=".gpx", full.names=TRUE )
df <- NULL

for (afile in in_files) {
  print(afile)
  
  foo <- XML::htmlTreeParse(afile , useInternalNodes = TRUE)
  t_coords <- XML::xpathSApply(doc = foo, path = "//trkpt", fun = XML::xmlAttrs)
  t_datetime <- XML::xpathSApply(doc = foo, path = "//trkpt/time", fun = XML::xmlValue)
  t_elevation <- XML::xpathSApply(doc = foo, path = "//trkpt/ele", fun = XML::xmlValue)
  ways <- XML::xpathSApply(doc = foo, path = "//wpt", fun = XML::xmlAttrs)
  wpt_type <- XML::xpathSApply(doc = foo, path = "//wpt/name", fun = XML::xmlValue)
  elevation <- XML::xpathSApply(doc = foo, path = "//wpt/ele", fun = XML::xmlValue)
  wy_datetime <- XML::xpathSApply(doc = foo, path = "//wpt/time", fun = XML::xmlValue)
  
  #   waypoints first
  foo <- data.frame(
    lat = as.numeric(ways["lat", ]),
    lon = as.numeric(ways["lon", ]),
    waypt = wpt_type,
    elevation = as.numeric(elevation),
    dt = lubridate::ymd_hms(wy_datetime)
  )
  
  df <- rbind(df, foo)
  
  #   Now track points
  foo <- data.frame(
    lat = as.numeric(t_coords["lat", ]),
    lon = as.numeric(t_coords["lon", ]),
    waypt = NA,
    elevation = as.numeric(t_elevation),
    dt = lubridate::ymd_hms(t_datetime)
  )
  
  df <- rbind(df, foo)
}

df <- df %>% 
  arrange(dt) %>% 
  mutate(Date=as.character(date(dt)))

```

```{r}

#   Make some plots

df %>% 
  filter(!is.na(waypt)) %>% 
  filter(!waypt=="glass") %>% 
  ggplot(aes(x=waypt)) +
  geom_bar() +
  labs(title="May 16 - June 26 Trash Collection",
       x="Item class")

df %>% 
  filter(!is.na(waypt)) %>% 
  filter(!waypt=="glass") %>% 
  filter(dt>lubridate::mdy("06/19/2024")) %>% 
  ggplot(aes(x=waypt)) +
  geom_bar() +
  labs(title="June 20 - September 30 2024 Trash Collection",
       x="Item class")

```

##        Map

```{r}

#   add geohash

# df$Geohash <- geohash::gh_encode(df$lon, df$lat, 12)
df$Geohash <- paste(df$lon, df$lat)

df_sf <- df %>%
  group_by(Date, Geohash) %>%
    summarize(Date=first(Date),
              lat=first(lat),
              lon=first(lon),
              dt=first(dt),
              elevation=first(elevation),
              num=n(),
              waypt=list(waypt)) %>%
  sf::st_as_sf(coords=c("lon","lat"), crs=googlecrs) %>% 
  ungroup()

#   Project to UTM

df_utm <- sf::st_transform(df_sf, crs=Seattle)

#   Some stats

df_utm %>% 
  sf::st_drop_geometry() %>% 
  group_by(Date) %>% 
    mutate(Dist=Eucl(lag(X, default=first(X)), X, 
                     lag(Y, default=first(Y)), Y)) %>% 
  ungroup() %>% 
  filter(Dist<500) %>% 
  ggplot(aes(x=Dist)) +
    geom_histogram()

#   Turn into a set of lines

l <- df_utm %>% 
  arrange(dt) %>% 
  group_by(Date) %>% 
  summarize(do_union=FALSE) %>% 
  sf::st_cast("LINESTRING")

#   Smooth

ls <- smoothr::smooth(sf::st_geometry(l), method = "ksmooth",
                      smoothness=10)

#   Plot the results

l %>% 
  ggplot() +
  geom_sf(aes(color="red")) +
  geom_sf(data=ls)

#   Make a prettier map of path

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(l) +
  tmap::tm_lines()

```


##    Work on spikes

```{r}

#   Find skip distances and plot

#   Project to UTM

df_utm <- sf::st_transform(df_sf, crs=Seattle)

#     Eliminate duplicate coordinate points that have no trash

df_utm <- df_utm %>%     
  arrange(dt) %>% 
  mutate(X=sf::st_coordinates(.)[,1],
         Y=sf::st_coordinates(.)[,2]) 

#   Turn into a set of lines

l <- df_utm %>% 
  group_by(Date) %>% 
  summarize(do_union=FALSE) %>% 
  sf::st_cast("LINESTRING")

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(l) +
  tmap::tm_lines()

df_utm2 <- delta_Dist(df_utm)

df_utm2 %>% 
  mutate(Delta_elevation=lag(elevation, default=0)-elevation) %>% 
  ggplot(aes(x=Delta_elevation, y=Delta_dist)) +
  geom_point()

#   Big deltas

Spike_pct <- 0.8

df_delta <- df_utm2 %>% 
  filter(Delta_dist>Spike_pct*New_dist) %>% 
  select(Delta_dist, Date, geometry) %>% 
  sf::st_as_sf()

df_utm3 <- df_utm2 %>% 
  group_by(Date) %>% 
    mutate(NewX=(lag(X, default=first(X)) + lead(X, default=last(X)))/2) %>% 
    mutate(NewY=(lag(Y, default=first(Y)) + lead(Y, default=last(Y)))/2) %>% 
  ungroup() %>% 
  arrange(dt)

for (i in 1:nrow(df_utm3)) {
  if (df_utm3[i,]$Delta_dist>Spike_pct*df_utm3[i,]$New_dist) {
    new_point <- sf::st_point(c(df_utm3[i,]$NewX, df_utm3[i,]$NewY)) %>% 
                 sf::st_sfc(crs = Seattle)
    df_utm3[i,]$geometry <- sf::st_sfc(sf::st_geometry(new_point))
  }
}

df_delta_after <- df_utm3 %>% 
  filter(Delta_dist>Spike_pct*New_dist) %>% 
  # filter(Delta_dist>30) %>% 
  select(Delta_dist, Date, geometry) %>% 
  sf::st_as_sf()

#   Turn into a set of lines

l <- df_utm3 %>% 
  group_by(Date) %>% 
  summarize(do_union=FALSE) %>% 
  sf::st_cast("LINESTRING")

#   Smooth

# ls <- smoothr::smooth(sf::st_geometry(l), method = "ksmooth",
#                       smoothness=10)
#   Make a prettier map of path

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(l) +
  tmap::tm_lines() +
# tmap::tm_shape(ls) +
#   tmap::tm_lines(col="blue") +
tmap::tm_shape(df_delta) +
  tmap::tm_dots(fill="red", size=0.5, popup.vars="Delta_dist")+
tmap::tm_shape(df_delta_after) +
  tmap::tm_dots(fill="green", size=0.5)



df_utm2 %>% sf::st_drop_geometry() %>% 
  filter(Distance_L>100) %>% 
  ggplot(aes(x=Distance_L)) +
  geom_histogram()

df_utm2 %>% sf::st_drop_geometry() %>% 
  filter(Delta_dist>100) %>%
  mutate(LogDelta=log(Delta_dist)) %>% 
  ggplot(aes(x=Delta_dist)) +
  # ggplot(aes(x=LogDelta)) +
  geom_histogram()

```

##    Kalman filtering

https://boazsobrado.com/blog/2017/08/29/implementing-a-kalman-filter-in-r-for-android-gps-measurements/

https://bookdown.org/probability/bookdown-demo/kalman-filters.html

https://maddevs.io/blog/reduce-gps-data-error-on-android-with-kalman-filter-and-accelerometer/

##    Apply Kalman filter

```{r}

#   Do some tests

foo <- df_utm3 %>% 
  mutate(X=sf::st_coordinates(.)[,1],
         Y=sf::st_coordinates(.)[,2]) %>% 
  sf::st_drop_geometry() %>% 
  select(Date, X, Y, num, Delta_dist, New_dist)

get_change <- function(df1, df2) {
  # df2 <- df2 %>% sf::st_drop_geometry() %>% select(X2, Y2)
  df1 <- df1 %>% sf::st_drop_geometry() %>% select(X, Y)
  foo <- cbind(df1, df2) %>% 
    mutate(Dx=abs(X-X2), Dy=abs(Y-Y2), Delta=Eucl(X, X2, Y, Y2))  %>% 
    select(Dx, Dy, Delta)
  return(foo)
}

collect <- NULL

for (i in c(0.2, 0.5, 1, 2, 4)) {
  for (j in c(0.2, 0.5, 1, 2, 4)) {
    Label <- paste(i,j, sep="_")
    foo_k <- foo %>% 
      group_by(Date) %>% 
        mutate(a=Kalman(X, Y, i*Delta_dist/num, j*New_dist*3)) %>% 
        mutate(X2=a$X2,
               Y2=a$Y2) %>% 
      ungroup() %>% 
      select(X2, Y2)
    foobar <- get_change(foo, foo_k) %>% 
      rename(!!sym(paste0(Label,"_Dx")) := Dx,
             !!sym(paste0(Label,"_Dy")) := Dy,
             !!sym(paste0(Label,"_Delta")) := Delta)
    
    collect <- bind_cols(collect, foobar)
    
  }
}
  
collect %>% 
  select(ends_with("Dx")) %>% 
  pivot_longer(cols=everything(), names_to="Name") %>% 
  mutate(class=factor(Name)) %>% 
  ggplot(aes(x=value)) +
    geom_histogram() +
    facet_wrap(~class)

collect %>% 
  select(ends_with("Dy")) %>% 
  pivot_longer(cols=everything(), names_to="Name") %>% 
  mutate(class=factor(Name)) %>% 
  ggplot(aes(x=value)) +
    geom_histogram() +
    facet_wrap(~class)

collect %>% 
  select(ends_with("Delta")) %>% 
  pivot_longer(cols=everything(), names_to="Name") %>% 
  mutate(class=factor(Name)) %>% 
  ggplot(aes(x=value)) +
    geom_histogram() +
    facet_wrap(~class)
########   end testing

#   run it 

df_kal <- df_utm3 %>% 
  mutate(X=sf::st_coordinates(.)[,1],
         Y=sf::st_coordinates(.)[,2]) %>% 
  sf::st_drop_geometry() %>% 
  group_by(Date) %>% 
    mutate(a=Kalman(X, Y, 4*Delta_dist/num, 0.5*New_dist)) %>% 
    mutate(X2=a$X2,
           Y2=a$Y2) %>% 
  ungroup()

df_kal <- df_kal %>% 
  sf::st_drop_geometry() %>% 
  sf::st_as_sf(coords=c("X2", "Y2"), crs=Seattle)

l2 <- df_kal %>% 
  group_by(Date) %>% 
  summarize(do_union=FALSE) %>% 
  sf::st_cast("LINESTRING")

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(l) +
  tmap::tm_lines() +
tmap::tm_shape(l2) +
  tmap::tm_lines(col="red") +
# tmap::tm_shape(ls) +
#   tmap::tm_lines(col="blue") +
tmap::tm_shape(df_utm3 %>% select(Date, elevation, num, waypt)) +
  tmap::tm_dots(fill="red", size=0.5) +
tmap::tm_shape(df_kal %>% select(Date, elevation, num, waypt)) +
  tmap::tm_dots(fill="green", size=0.5)
```




