---
title: "Read data and clean up"
author: "Alan Jackson"
format: html
editor: source
---

##    Read files, project, and post on map

Read files in, and split into path and collection points
Run some simple stats on collection points
Project the data
Create a 25 m corridor around the path
Are there collection points outside that corridor?
If so create 10 meter circles around them, and do a union with corridor
Create a 5 m grid in the area
Using 20 m mean filter, populate grid
Contour and plot

```{r}

# setup

library(tidyverse)

path <- "/home/ajackson/Dropbox/Rprojects/TrashTracker/data/"


```

##        Read in all the files

Read in files
split
generate stats

```{r}

#   Read in the files

gpx_parsed <- list.files( path=paste0(path), pattern=".gpx", full.names=TRUE ) %>% 
  map_dfr( XML::htmlTreeParse(. , useInternalNodes = TRUE))

coords <- XML::xpathSApply(doc = gpx_parsed, path = "//trkpt", fun = XML::xmlAttrs)
ways <- XML::xpathSApply(doc = gpx_parsed, path = "//wpt", fun = XML::xmlAttrs)
wpt_type <- XML::xpathSApply(doc = gpx_parsed, path = "//wpt/name", fun = XML::xmlValue)
elevation <- XML::xpathSApply(doc = gpx_parsed, path = "//wpt/ele", fun = XML::xmlValue)
# elevation <- XML::xpathSApply(doc = gpx_parsed, path = "//trkpt/ele", fun = XML::xmlValue)

df <- data.frame(
  lat = as.numeric(ways["lat", ]),
  lon = as.numeric(ways["lon", ]),
  waypt = wpt_type,
  elevation = as.numeric(elevation)
)

```
